---
title: "Lab 2"
author: "Naveen Mathew Nathan S."
output: html_document
---

#### 1. Changing the author field and file name.

##### (a) Change the `author:` field on the Rmd document from Your Name Here to your own name.

##### (b) Rename this file to "lab02_YourHameHere.Rmd", where YourNameHere is changed to your own name.

#### 2. Data practice

In class we imported the survey data using the `read.table()` function.  (Note: on Homework 1 you'll be asked to use the `read.csv()` function instead.)

This is the code we used:

```{r}
survey <- read.table("http://www.andrew.cmu.edu/user/achoulde/94842/data/survey_data.csv", header=TRUE, sep=",") 
```

##### (a) How many survey respondents are from MISM or Other?

```{r}
mean(survey$Program == "MSIM" | survey$Program == "Other")
```

##### (b) What % of survey respondents are from PPM?

```{r}
mean(survey$Program == "PPM")
```

#### 3. Index practice

##### (a) Use $ notation to pull the OperatingSystem column from the survey data

```{r}
print(survey$OperatingSystem)
```

##### (b) Do the same thing with [,] notation, referring to OperatingSystem by name

```{r}
print(survey[, "OperatingSystem"])
```

##### (c) Repeat part (b), this time referring to OperatingSystem by column number

```{r}
# Identifying the index
print(which(colnames(survey) == "OperatingSystem"))
print(survey[, 4])
```



