---
title: Tutorial 3 - Manipulating data, lists, functions; if-else statements
font-family: Gill Sans
autosize: false
---
### Content

- More on data frames
- Lists
- Writing functions in R
- If-else statements


###More on data frames
```{r}
library(MASS)
head(Cars93, 3)
```

###Adding a column: `transform()` function
- `transform()` returns a new data frame with columns modified or added as specified by the function call

```{r}
Cars93.metric <- transform(Cars93, 
                           KMPL.city = 0.425 * MPG.city, 
                           KMPL.highway = 0.425 * MPG.highway)
tail(names(Cars93.metric))
```

- Our data frame has two new columns, giving the fuel consumption in km/l

###Another approach
```{r}
# Add a new column called KMPL.city.2
Cars93.metric$KMPL.city.2 <- 0.425 * Cars93$MPG.city
tail(names(Cars93.metric))
```

- Let's check that both approaches did the same thing

```{r}
identical(Cars93.metric$KMPL.city, Cars93.metric$KMPL.city.2)
```


###Changing levels of a factor
```{r}
manufacturer <- Cars93$Manufacturer
head(manufacturer, 10)
```

We'll use the `mapvalues(x, from, to)` function from the `plyr` library.

```{r}
library(plyr)

# Map Chevrolet, Pontiac and Buick to GM
manufacturer.combined <- mapvalues(manufacturer, 
                                   from = c("Chevrolet", "Pontiac", "Buick"), 
                                   to = rep("GM", 3))

head(manufacturer.combined, 10)
```

###Another example
- A lot of data comes with integer encodings of levels

- You may want to convert the integers to more meaningful values for the purpose of your analysis

- Let's pretend that in the class survey 'Program' was coded as an integer with 1 = MISM, 2 = Other, 3 = PPM

```{r}
survey <- read.table("http://www.andrew.cmu.edu/user/achoulde/94842/data/survey_data.csv", header=TRUE, sep=",") 
survey <- transform(survey, Program=as.numeric(Program))
head(survey)
```

###Example continued

- Here's how we would get back the program codings using the `transform()`, `as.factor()` and `mapvalues()` functions

```{r}
survey <- transform(survey, 
                    Program = as.factor(mapvalues(Program, 
                                                  c(1, 2, 3), 
                                                  c("MISM", "Other", "PPM")))
                    )
head(survey)
```


###Some more data frame summaries: `table()` function

- Let's revisit the Cars93 dataset

- The `table()` function builds **contingency tables** showing counts at each combination of factor levels

```{r}
table(Cars93$AirBags)
```
###
```{r}
table(Cars93$Origin)
table(Cars93$AirBags, Cars93$Origin)
```

- Looks like US and non-US cars had about the same distribution of AirBag types

- Later in the class we'll learn how to do a hypothesis tests on this kind of data

###Alternative syntax

- When `table()` is supplied a data frame, it produces contingency tables for all combinations of factors 

```{r}
head(Cars93[c("AirBags", "Origin")], 3)
table(Cars93[c("AirBags", "Origin")])
```

###Basics of lists

> A list is a **data structure** that can be used to store **different kinds** of data

- Recall: a vector is a data structure for storing *similar kinds of data*

- To better understand the difference, consider the following example.

```{r}
my.vector.1 <- c("Michael", 165, TRUE) # (name, weight, is.male)
my.vector.1 
typeof(my.vector.1)  # All the elements are now character strings!
```

###Lists vs. vectors

```{r}
my.vector.2 <- c(FALSE, TRUE, 27) # (is.male, is.citizen, age)
typeof(my.vector.2)
```

- Vectors expect elements to be all of the same type (e.g., `Boolean`, `numeric`, `character`)

- When data of different types are put into a vector, the R converts everything to a common type

###Lists

- To store data of different types in the same object, we use lists

- Simple way to build lists: use `list()` function

```{r}
my.list <- list("Michael", 165, TRUE)
my.list
sapply(my.list, typeof)
```

###Named elements
```{r}
patient.1 <- list(name="Michael", weight=165, is.male=TRUE)
patient.1
```

###Referencing elements of a list (similar to data frames)
```{r}
patient.1$name # Get "name" element (returns a string)
patient.1[["name"]] # Get "name" element (returns a string)
patient.1["name"] # Get "name" slice (returns a sub-list)
c(typeof(patient.1$name), typeof(patient.1["name"]))
```


###Functions
- We have used a lot of built-in functions: `mean()`, `subset()`, `plot()`, `read.table()`...

- An important part of programming and data analysis is to write custom functions

- Functions help make code **modular**

- Functions make debugging easier

- Remember: this entire class is about applying *functions* to *data*

###What is a function?

> A function is a machine that turns **input objects** (arguments) into an **output object** (return value) according to a definite rule.

- Let's look at a really simple function

```{r}
addOne <- function(x) {
  x + 1
}
```

- `x` is the **argument** or **input**

- The function **output** is the input `x` incremented by 1

```{r}
addOne(12)
```

###More interesting example

- Here's a function that returns a % given a numerator, denominator, and desired number of decimal values

```{r}
calculatePercentage <- function(x, y, d) {
  decimal <- x / y  # Calculate decimal value
  round(100 * decimal, d)  # Convert to % and round to d digits
}

calculatePercentage(27, 80, 1)
```

- If you're calculating several %'s for your report, you should use this kind of function instead of repeatedly copying and pasting code

###Function returning a list
- Here's a function that takes a person's full name (FirstName LastName), weight in lb and height in inches and converts it into a list with the person's first name, person's last name, weight in kg, height in m, and BMI.

```{r}
createPatientRecord <- function(full.name, weight, height) {
  name.list <- strsplit(full.name, split=" ")[[1]]
  first.name <- name.list[1]
  last.name <- name.list[2]
  weight.in.kg <- weight / 2.2
  height.in.m <- height * 0.0254
  bmi <- weight.in.kg / (height.in.m ^ 2)
  list(first.name=first.name, last.name=last.name, weight=weight.in.kg, height=height.in.m,
       bmi=bmi)
}
```
 
###Trying out the function
```{r}
createPatientRecord("Michael Smith", 185, 12 * 6 + 1)
```

###Another example: 3 number summary
- Calculate mean, median and standard deviation
```{r}
threeNumberSummary <- function(x) {
  c(mean=mean(x), median=median(x), sd=sd(x))
}
x <- rnorm(100, mean=5, sd=2) # Vector of 100 normals with mean 5 and sd 2
threeNumberSummary(x)
```


###If-else statements

- Oftentimes we want our code to have different effects depending on the features of the input

- Example: Calculating a student's letter grade
  - If grade >= 90, assign A
  - Otherwise, if grade >= 80, assign B
  - Otherwise, if grade >= 70, assign C
  - In all other cases, assign F

- To code this up, we use if-else statements

###If-else Example: Letter grades
```{r}
calculateLetterGrade <- function(x) {
  if(x >= 90) {
    grade <- "A"
  } else if(x >= 80) {
    grade <- "B"
  } else if(x >= 70) {
    grade <- "C"
  } else {
    grade <- "F"
  }
  grade
}

course.grades <- c(92, 78, 87, 91, 62)
sapply(course.grades, FUN=calculateLetterGrade)
```

###`return()`

- In the previous examples we specified the output simply by writing the output variable as the last line of the function

- More explicitly, we can use the `return()` function

```{r}
addOne <- function(x) {
  return(x + 1)
}

addOne(12)
```

- We will generally avoid the `return()` function, but you can use it if necessary or if it makes writing a particular function easier.

### Next 


- Complete [Lab 3](http://isle.heinz.cmu.edu/94-842/lab03/)
