---
title: Tutorial 4 - Basic cleaning, loops, apply functions

font-family: Gill Sans
---

### Contents

- A common data cleaning task
- If-else statements
- For/while loops to iterate over data
- `apply()`, `lapply()`, `sapply()`, `tapply()`
- `with()` to specify scope

###A common problem

- One of the most common problems you'll encounter when importing manually-entered data is inconsistent data types within columns

- For a simple example, let's look at `TVhours` column in a messy version of the survey data from Tutorial 2

```{r}
survey.messy <- read.csv("http://www.andrew.cmu.edu/user/achoulde/94842/data/survey_messy.csv", header=TRUE)
survey.messy$TVhours
```

###What's happening?
```{r}
str(survey.messy)
```
- Several of the entries have non-numeric values in them (they contain strings)

- As a result, `TVhours` is being imported as `factor`

###Attempt at a fix
- What if we just try to cast it back to numeric?

```{r}
tv.hours.messy <- survey.messy$TVhours
tv.hours.messy
as.numeric(tv.hours.messy)
```

###That didn't work...
```{r, eval = FALSE}
tv.hours.messy
as.numeric(tv.hours.messy)
```
```{r, echo = FALSE}
# Show this output, but not this code
head(tv.hours.messy, 40)
head(as.numeric(tv.hours.messy), 40)
```



- This just converted all the values into the integer-coded levels of the factor

- Not what we wanted!

###Something that does work

- Consider the following simple example

```{r}
num.vec <- c(3.1, 2.5)
as.factor(num.vec)
as.numeric(as.factor(num.vec))
as.numeric(as.character(as.factor(num.vec)))
```

> If we take a number that's being coded as a factor and *first* turn it into a `character` string, *then* converting the string to a numeric gets back the number

###Back to the corrupted TVhours column

```{r}
as.character(tv.hours.messy)
as.numeric(as.character(tv.hours.messy))
typeof(as.numeric(as.character(tv.hours.messy)))  # Success!! (Almost...)
```

###A small improvement

- All the corrupted cells now appear as `NA`, which is R's missing indicator

- We can do a little better by cleaning up the vector once we get it to `character` form

```{r}
tv.hours.strings <- as.character(tv.hours.messy)
tv.hours.strings
```

###Deleting non-numeric (or .) characters 
```{r}
tv.hours.strings
# Use gsub() to replace everything except digits and '.' with a blank ""
gsub("[^0-9.]", "", tv.hours.strings) 
```

###The final product
```{r}
tv.hours.messy[1:30]
tv.hours.clean <- as.numeric(gsub("[^0-9.]", "", tv.hours.strings))
tv.hours.clean
```
- As a last step, we should go through and figure out if any of the `NA` values should really be `0`.  
  - This step is not shown here.

###Rebuilding our data
```{r}
survey <- transform(survey.messy, TVhours = tv.hours.clean)
str(survey)
```

- **Success!**

###A different approach

- We can also handle this problem by setting `stringsAsFactors = FALSE` when importing our data.

```{r}
survey.messy <- read.csv("http://www.andrew.cmu.edu/user/achoulde/94842/data/survey_messy.csv", header=TRUE, stringsAsFactors=FALSE)
str(survey.messy)
```

- Now everything is a `character` instead of a `factor`

###One-line cleanup 

- Let's clean up the `TVhours` column and cast it to numeric all in one command

```{r}
survey <- transform(survey.messy, 
                    TVhours = as.numeric(gsub("[^0-9.]", "", TVhours)))
str(survey)
```

###What about all those other `character` variables?

```{r}
table(survey[["Program"]])
table(as.factor(survey[["Program"]]))
```

- Having factors coded as characters may be OK for many parts of our analysis

###To be safe, let's fix things
```{r}
# Figure out which columns are coded as characters
chr.indexes <- sapply(survey, FUN = is.character)
chr.indexes
# Re-code all of the character columns to factors
survey[chr.indexes] <- lapply(survey[chr.indexes], FUN = as.factor)
```

###Here's the outcome
```{r}
str(survey)
```

- **Success!**

###Another common problem

- On Homework 2 you'll learn how to wrangle with another common problem

- When data is entered manually, misspellings and case changes are very common

- E.g., a column showing life support mechanism may look like,

```{r}
life.support <- as.factor(c("dialysis", "Ventilation", "Dialysis", "dialysis", "none", "None", "nnone", "dyalysis", "dialysis", "ventilation", "none"))
summary(life.support)
```
###
```{r}
summary(life.support)
```

- This factor has 8 levels even though it should have 3 (dialysis, ventilation, none)

- We can fix many of the typos by running spellcheck in Excel before importing data, or by changing the values on a case-by-case basis later

- There's a faster way to fix just the capitalization issue (this is an exercise on Homework 2)


###What are all these [l/s/t/]apply() functions?

- These are all efficient ways of applying a function to margins of an array or elements of a list

- Before we talk about the details of `apply()` and its relatives, we should first understand loops

- **loops** are ways of iterating over data

- The `apply()` functions can be thought of as good *alternatives* to loops 

###For loops: a pair of examples
```{r}
for(i in 1:4) {
  print(i)
}

phrase <- "Good Night, "
for(word in c("and", "Good", "Luck")) {
  phrase <- paste(phrase, word)
  print(phrase)
}
```

###For loops: syntax

> A **for loop** executes a chunk of code for every value of an **index variable** in an **index set**

- The basic syntax takes the form

```{r, eval=FALSE}
# for(index.variable in index.set) {
#   code to be repeated at every value of index.variable
# }
```

- The index set is often a vector of integers, but can be more general

###Example
```{r}
index.set <- list(name="Michael", weight=185, is.male=TRUE) # a list
for(i in index.set) {
  print(c(i, typeof(i)))
}
```

###Example: Calculate sum of each column
```{r}
fake.data <- matrix(rnorm(500), ncol=5) # create fake 100 x 5 data set
head(fake.data,2) # print first two rows

col.sums <- numeric(ncol(fake.data)) # variable to store running column sums
for(i in 1:nrow(fake.data)) {
  col.sums <- col.sums + fake.data[i,] # add ith observation to the sum
}
col.sums

colSums(fake.data) # A better approach (see also colMeans())
```

###while loops

- **while loops** repeat a chunk of code while the specified condition remains true

```{r, eval=FALSE}
day <- 1
num.days <- 365
while(day <= num.days) {
  day <- day + 1
}
```

- We won't really be using while loops in this class

- Just be aware that they exist, and that they may become useful to you at some point in your analytics career

###The various apply() functions

Command | Description
--------|------------
`apply(X, MARGIN, FUN)` | Obtain a vector/array/list by applying `FUN` along the specified `MARGIN` of an array or matrix `X`
`lapply(X, FUN)` | Obtain a list by applying `FUN` to the elements of a list `X`
`sapply(X, FUN)` | Simplified version of `lapply`.  Returns a vector/array instead of list.
`tapply(X, INDEX, FUN)` | Obtain a table by applying `FUN` to each combination of the factors given in `INDEX`  

- These functions are (good!) alternatives to loops

- They are typically *more efficient* than loops (often run considerably faster on large data sets)

- Take practice to get used to, but make analysis easier to debug and less prone to error when used effectively

- You can always type `example(function)` to get code examples (E.g., `example(apply)`)

###Example: apply()
```{r}
colMeans(fake.data)
apply(fake.data, MARGIN=2, FUN=mean) # MARGIN = 1 for rows, 2 for columns
# Function that calculates proportion of vector indexes that are > 0
propPositive <- function(x) mean(x > 0)
apply(fake.data, MARGIN=2, FUN=propPositive) 
```

###Example: lapply(), sapply()
```{r}
lapply(survey, is.factor) # Returns a list
sapply(survey, FUN = is.factor) # Returns a vector with named elements
```

###Example: apply(), lapply(), sapply()
```{r}
apply(cars, 2, FUN=mean) # Data frames are arrays
lapply(cars, FUN=mean) # Data frames are also lists
sapply(cars, FUN=mean) # sapply() is just simplified lapply()
```


###Example: tapply()

- Think of `tapply()` as a generalized form of the `table()` function

```{r}
library(MASS)
# Get a count table, data broken down by Origin and DriveTrain
table(Cars93$Origin, Cars93$DriveTrain)

# Calculate average MPG.City, broken down by Origin and Drivetrain
tapply(Cars93$MPG.city, INDEX = Cars93[c("Origin", "DriveTrain")], FUN=mean)
```

###Example: tapply()

- Let's get the average horsepower by car `Origin` and `Type`

```{r}
tapply(Cars93[["Horsepower"]], INDEX = Cars93[c("Origin", "Type")], FUN=mean)
```

- What's that `NA` doing there?

```{r}
any(Cars93$Origin == "non-USA" & Cars93$Type == "Large")
```

- None of the non-USA manufacturers produced Large cars!

###Example: using tapply() to mimic table()

- Here's how one can use `tapply()` to produce the same output as the `table()` function

```{r}
library(MASS)
# Get a count table, data broken down by Origin and DriveTrain
table(Cars93$Origin, Cars93$DriveTrain)

# This one may take a moment to figure out...
tapply(rep(1, nrow(Cars93)), INDEX = Cars93[c("Origin", "DriveTrain")], FUN=sum) 
```

###with()

- Thus far we've repeatedly typed out the data frame name when referencing its columns

- This is because the data variables don't exist in our working environment

- Using **with**`(data, expr)` lets us specify that the code in `expr` should be evaluated in an environment that contains the elements of `data` as variables

```{r}
with(Cars93, table(Origin, Type))
```

###Example: with()

```{r}
any(Cars93$Origin == "non-USA" & Cars93$Type == "Large")
with(Cars93, any(Origin == "non-USA" & Type == "Large")) # Same effect!
```

```{r}
with(Cars93, tapply(Horsepower, INDEX = list(Origin, Type), FUN=mean))
```

- Using with() makes code simpler, easier to read, and easier to debug

### Next 
- Complete [**Lab 4**](http://isle.heinz.cmu.edu/94-842/lab04/) 