---
title: "Example: GLM"
output: pdf_document
---

## Example

```{r}
snoreheart <- read.table("snoreheart.txt", header = T)
snorefit <- glm(cbind(Disease, NoDisease) ~ Snoring, data = snoreheart, family = "binomial")
print(summary(snorefit))
print(deviance(snorefit))
print(df.residual(snorefit))
print(1 - pchisq(deviance(snorefit), df.residual(snorefit)))
```

We cannot reject the null hypothesis that the model is sufficient.

```{r}
new_scores <- c(0, 2, 4, 6) # Last score was 5 in the original data set
snorefit2 <- glm(cbind(Disease, NoDisease) ~ new_scores, data = snoreheart, family = "binomial")
print(deviance(snorefit2))
print(df.residual(snorefit2))
print(1 - pchisq(deviance(snorefit2), df.residual(snorefit2)))
```

We reject the null hypothesis that the model is sufficient.

## Nested model comparison

Consider GLMs $M_0$ and $M_1$ with $M_0$ nested within $M_1$ (every distribution in $M_0$ is also in $M_1$, but not vice versa)

Therefore, $\mu$ is more restricted under $M_0$ than under $M_1$.

Let $\hat\mu_0$ be MLE of $M_0$ and $\hat\mu_1$ be MLE of $M_1$.

$H_0$: $M_0$ true vs $H_1$: $M_1$ true, but not $M_0$

LRT chi-squared statistic: $-2(L(\hat\mu_0; y)-L(\hat\mu_1; y)) = -2(L(\hat\mu_0;y)-L(y;y))-(-2(L(\hat\mu_1;y)-L(y;y))) = D(y; \hat\mu_0)-D(y; \hat\mu_1)$

Deviance decreases as the model becomes larger. Similar to sum of squares in linear regression.

df = effective number of parameters in $M_1$ - effective number of parameters in $M_0$

Reject $H_0$ if $G^2(M_0|M_1)=D(y; \hat\mu_0)-D(y; \hat\mu_1) > \chi^2_{df}(\alpha)$

Assuming $M_1$ is not too close to the saturated model, the approximation holds true even if the counts are less than 5.

## Profile likelihood CIs

If Wald CIs are not reliable, 

$M_0(\beta_0) =$ same model, except $\beta$ is fixed at $\beta_0$, df = 1

Then LRT tests $H_0: \beta = \beta_0; H_a: \beta \neq \beta_0$

Looks like this should be repeated for each value of $\beta_0$, but there are computational simplifications that can be used with just one $\beta_0$

## Example

```{r}
horseshoe <- read.table("horseshoe.txt", header = T)
model <- glm(satell ~ color + spine + width + weight, family = poisson, data = horseshoe)
summary(model)
```