---
title: "Matched Pairs"
output: pdf_document
---

## Generalized matched case-control

- More than 1 X variable on which strata are created ($X_1, ..., X_p$)
- More than 1 control for each case (one-to-many matching) - matched set instead of matched pair (not allowed to match based on $X_1, ..., X_p$)

Let n = total number of matched sets, $T_i$ = total number of subjects in ith matched set, $Y_{it}$ = response Y for subject t in matched set i: will be 1 for only one t for a given i (only one control in each set), $x_{hit}$ = value of $X_h$ for subject t in matched set i

h = 1, ..., p; i = 1, ..., n; t = 1, ..., $T_i$

## Model

Logit link: $logit(P(Y_{it}=1)) = \alpha_i + \beta_1 x_{1it} + ... + \beta_p x_{pit} = \alpha_i + x_{it}^T \beta$ with $x_{it}^T = (x_{1it}, ..., x_{pit})$; conditioned on $S_i = \sum_{t=1}^{T_i} {Y_{it}} = 1$

$P(Y_{i\tau}=1|S_i=1) = \frac{e^{x_{i\tau}} \beta}{\sum_{t=1}^{T_i}e^{x_{i\tau}} \beta}$

Let $\tau_i$ be is the index such that $y_{i\tau_i}=1$, then $\prod_{i=1}^n \frac{e^{x_{i\tau}} \beta}{\sum_{t=1}^{T_i}e^{x_{i\tau}} \beta}$ is conditional likelihood for $\beta$. Using it is conditional logistic regression.

## Example

```{r}
library(faraway)
data(amlxray)
head(amlxray)
```

```{r}
subset(amlxray, downs=="yes")
```

If `downs` is used as predictor, MLE will not exist (quasi complete separation)

Restricting analysis to individuals (matched sets) who don't have Down's syndrome

```{r}
nonDown <- subset(amlxray, !(ID %in% subset(amlxray, downs=="yes")$ID))
table(nonDown$ID)
```

Only one-to-one or one-to-two matching observed in the data set.

```{r}
library(survival)
mod <- clogit(disease ~ Sex + Mray + Fray + CnRay + strata(ID), data=nonDown)
summary(mod)
```

- CnRay.L: linear
- CnRay.Q: quadratic (orthogonalized with respect to linear and intercept terms)
- CnRay.C: cubic (orthogonalized with respect to quadratic, linear and intercept terms)

exp(coef) is odds ratio

Table also shows transformed Wald 95% CI of the odds ratio (exp(coef)).

To get a single p-value for CnRay variable:

```{r}
drop1(mod, test="Chisq")
```

Difficult to interpret CnRay.L, CnRay.Q, CnRay.C as they are for the orthogonalized linear, quadratic and cubic terms of CnRay.