---
title: "Other link functions"
output: pdf_document
---

Logit or probit: interchanging $n_i-Y_i$ and $Y_i$ gives new coefficient = -1 * old coefficient.

Curve converges equally quickly to 1 as to 0.


# Complementary log-log

$g(\pi) = log(-log(1-\pi))$

$\pi(\eta) = 1-e^{-e^\eta}$ is a cumulative distribution function of Gumbel distribution (useful extreme value distribution)

$\beta > 0 \implies$ faster convergence to 1 than to 0 - useful in toxicology studies. Poisons kill lower proportion at lower doses, but kill almost all at high doses.

$(1-\pi(x_2)) = (1-\pi(x_1))^{e^{\beta (x_2-x_1)}}$ looks similar to propotional hazards

- Does not have symmetry property. Depends on $Y_i$ and $n_i-Y-i$ - interchanging it gives model with log-log link, which is different model compared to complementary log-log link.

## Example

```{r}
beetle <- data.frame(logdose = c(1.6907, 1.7242, 1.7552, 1.7842, 1.8113,
1.8369, 1.8610, 1.8839),
number = c(59, 60, 62, 56, 63, 59, 62, 60),
killed = c( 6, 13, 18, 28, 52, 53, 61, 60))
plot(killed/number ~ logdose, data=beetle, ylim=c(0,1))
```

### Trying all links

```{r}
mod.logit <- glm(cbind(killed, number - killed) ~ logdose, data = beetle, family = binomial)
summary(mod.logit)
mod.probit <- glm(cbind(killed, number - killed) ~ logdose, data = beetle, family = binomial(link = probit))
summary(mod.probit)
```

AIC can be used to compare logit and probit link models because the outcome is the same and the model is still binomial.

```{r}
mod.cloglog <- glm(cbind(killed, number - killed) ~ logdose, data = beetle, family = binomial(link = cloglog))
summary(mod.cloglog)
```

Looks like the better model for prediction based on AIC.

**Deviance based tests**

```{r}
1- pchisq(deviance(mod.logit), df.residual(mod.logit))
1- pchisq(deviance(mod.probit), df.residual(mod.probit))
1- pchisq(deviance(mod.cloglog), df.residual(mod.cloglog))
```

Validity may be questionable as there are few small counts.

**Plotting the curves**

```{r}
plot(killed/number ~ logdose, data=beetle, ylim=c(0,1))
x <- beetle$logdose
curve(predict(mod.logit, data.frame(logdose=x), type="response"),
add=TRUE, lty=1, col="black")
curve(predict(mod.probit, data.frame(logdose=x), type="response"),
add=TRUE, lty=2, col="blue", lwd=2)
curve(predict(mod.cloglog, data.frame(logdose=x), type="response"),
add=TRUE, lty=3, col="red", lwd=3)
legend("bottomright", c("Logistic","Probit","Complementary Log-Log"),
lty=1:3, col=c("black","blue","red"), lwd=1:3)
```


# Multinomial response

Outcome variable Y with J possible categories

Can be:

- nominal (there is no natural order)
- ordinal

Intercept may be present, but it is not explicitly added to x.

$\pi_j(x_) = P(Y\ in\ category\ j | x); j = 1, 2, ..., J\ st \sum_{j=1}^J\pi_j(x)=1$ for any x