---
title: "Example: GLM"
output: pdf_document
---

## Example

```{r}
horseshoe <- read.table("horseshoe.txt", header = T)
model <- glm(satell ~ color + spine + width + weight, family = poisson, data = horseshoe)
summary(model)
```

```{r}
pchisq(model$deviance, model$df.residual)
```

**Note:** Looks like there is lack of fit in the data. Therefore estimates and levels of significance may by incorrect.

**Variance - covariance matrix**

```{r}
vcov(model)
```

**Rebuilding model**

```{r}
model0 <- glm(satell ~ color + weight, family = poisson, data = horseshoe)
summary(model0)
```

**Likelihood ratio test**

```{r}
anova(model0, model, test = "Chisq")
```

**Choosing variable to drop (LRT)**

```{r}
drop1(model, test = "Chisq")
```

Chisq test can be used as an alternative to Wald test. Compare `LRT` with chisq with 1 df.

$H_0:$ full model is required (variable is significant) vs $H_1:$ reduced model is sufficient (variable is insignificant).

**Score test**

```{r}
anova(model0, model, test = "Rao")
```

**Choosing variable to drop (Score)**

```{r}
drop1(model, test = "Rao")
```

**Profile likelihood CI**

```{r}
confint(model)
```

# Residuals

Used to examine lack of fit: patterns of departure from model and "outliers".

- Raw residuals: $y_i - \hat\mu_i$
- Pearson residuals: $\frac{y_i - \hat\mu_i}{\sqrt{\nu(\hat\mu_i)}}$. Squaring and adding will give generalized Pearson chi square.
- Eg: for Poisson: $\frac{y_i - \hat\mu_i}{\sqrt{\hat\mu_i}}$
- Deviance residuals: $d_i = -2(L(\hat\mu_i; y_i)-L(y_i;y_i))$ is always negative; $D(y, \hat\mu) = \sum_{i=1}^{N}d_i$
- ith deviance residual: $sign(y_i-\hat\mu_i)\sqrt{d_i}$; sum of squares of deviance residuals gives deviance
- Neither Pearson nor deviance residuals are standardized. Variance tends to be less than 1
- Standardized residuals: $r_i = \frac{y_i-\hat\mu_i}{\sqrt{\nu(\hat\mu_i)(1-\hat h_i)}} = \frac{e_i}{\sqrt{1-\hat h_i}}$; $h_i$ is the ith leverage = ith diagonal element of leverage matrix: $\hat H = \hat W^{1/2} X (X^T\hat WX)^{-1} X^T \hat W^{1/2} = \hat W^{1/2} X I^{-1} X^T \hat W^{1/2}$

Leverage is a measure of potential influence.

## Residuals in example

**Raw**

```{r}
plot(residuals(model, type = "response"))
```

Raw residuals look skewed.

**Pearson residuals**

```{r}
plot(residuals(model, type = "pearson"))
(X.2 <- sum(residuals(model, type = "pearson")^2))
model$deviance
```

We expect 5% to be outside $\pm 2$ range. Looks like there is some issue with the model.

**Deviance residuals**

```{r}
plot(residuals(model, type = "deviance"))
(X.2 <- sum(residuals(model, type = "deviance")^2))
model$deviance
```