---
title: "Ordinal response"
output: pdf_document
---

## Example

```{r}
happiness <- read.table("happiness.txt", header = T)
library(VGAM)
mod <- vglm(happy ~ trauma + race, data = happiness, family = propodds(reverse = F)) # Reverse changes sign of beta
summary(mod)
```

Coefficients of trauma and race are **negative**. Therefore there is **positive** relationship between trauma and response variable. But happiness is in reverse order (1 = very happy, 2 = pretty happy, 3 = not so happy). Therefore, higher trauma is related to lower levels of happiness.

- Odds of "very happy" or "pretty happy" for white relative to black (trauma held fixed): $e^{2.03612} = 7.660827$ (white is 0, black is 1, therefore difference is -1)
- Wald 95% CI: $e^{2.03612 \pm 1.96 \times 0.69113} = (1.976851, 29.687759)$

### LRT for race effect (controlling for trauma)

```{r}
modred <- vglm(happy ~ trauma, data = happiness, family = propodds(reverse = F))
llrts <- deviance(modred) - deviance(mod)
llrts.df <- df.residual(modred) - df.residual(mod)
1 - pchisq(llrts, llrts.df)
```

Race effect is quite significant.

### Estimating probabilities

```{r}
predict(mod, newdata = data.frame(trauma = mean(happiness$trauma), race = 0),
        type = "response")
```

### Probabilities vs trauma

```{r}
x = happiness$trauma
curve(predict(mod, data.frame(trauma=x,race=0), type="response")[,1],
xlab="Trauma", ylab="Probability",
xlim=range(happiness$trauma), ylim=c(0,1))
curve(predict(mod, data.frame(trauma=x,race=0), type="response")[,2],
add=TRUE, col="red")
curve(predict(mod, data.frame(trauma=x,race=0), type="response")[,3],
add=TRUE, col="blue")
curve(predict(mod, data.frame(trauma=x,race=1), type="response")[,1],
add=TRUE, lty=2)
curve(predict(mod, data.frame(trauma=x,race=1), type="response")[,2],
add=TRUE, col="red", lty=2)
curve(predict(mod, data.frame(trauma=x,race=1), type="response")[,3],
add=TRUE, col="blue", lty=2)
legend("topright", c("white","black"), lty=1:2)
legend("topleft", c("Very Happy","Pretty Happy","Not Too Happy"), lty=1,
col=c("black","red","blue"))
```

### Fitting non-proportional model

```{r}
modnotprop <- vglm(happy ~ trauma + race, data = happiness, family = cumulative(parallel = F))
llrts <- deviance(mod) - deviance(modnotprop)
llrts.df <- df.residual(mod) - df.residual(modnotprop)
1 - pchisq(llrts, llrts.df)
```

Reduced model is sufficient.

```{r}
library(MASS)
mod <- polr(factor(happy) ~ trauma + race, data = happiness)
summary(mod)
```

Assumes different parameterization, therefore the opposite signs of beta.

Standard errors are different because this uses observed information rather than expected Fisher information.

If happy is recoded as (3 = very happy, 2 = , 1 = not so happy), the signs of intercepts and coefficients will reverse, and intercepts will be rearranged.