---
title: "Other models: Ordinal response"
output: pdf_document
---

# Latent variable formulation

Y derived from latent continuous variable $Y^*$. Eg: $Y=j if \alpha_{j-1}<Y^{*}\le \alpha_j$

Suppose $Y^*=\beta^Tx + \epsilon$, where $\epsilon$ has known cdf G. Adding intercept to Y\* gives cut points at $\alpha + intercept$, which is the same model. Alternatively we can fix $\alpha_1=0$. Positive slope implies positive relationship between Y and X. Y\* independent implies Y is independent.

$P(Y \le j | x) = P(Y^*\le\alpha_j|x)=P(Y^*-\beta^Tx|x)=G() \implies G^{-1}(P(Y^*-\beta^Tx|x)) = \alpha_j-\beta^Tx$

Similar to proportional odds model, but G can be different. G = logit gives same model as proportional odds, but sign of $\beta$ is reversed.

Cumulative probit model: $G = \phi$

Proportional hazards: G^{-1}(p) = ln(-ln(1-p)) [used in survival analysis when there is grouped survival times]

## Example

```{r}
library(VGAM)
happiness <- read.table("happiness.txt", header = T)
mod.logit <- vglm(happy ~ trauma + race, family = cumulative(parallel = T), data = happiness)
summary(mod.logit)
mod.probit <- vglm(happy ~ trauma + race, family = cumulative(parallel = T, link = "probit"), data = happiness)
summary(mod.probit)
```

Saturated models are the same for logit and probit. Therefore, we can compare deviances.

Another way of fitting:

```{r}
# Logit link
library(MASS)
mod.logit2 <- polr(factor(happy) ~ trauma + race, data = happiness)
summary(mod.logit2)
```

Unlike `vglm`, `polr` uses $\alpha_j - \beta^Tx$. Positive $\beta$ indicates positive relationship between Y and x.

- Threshold between level 1 and 2: -0.5181
- Threshold between level 2 and 3: 3.4006

```{r}
# Fitting probit
mod.probit2 <- polr(factor(happy) ~ trauma + race, data = happiness, method = "probit")
```