---
title: "Tests for independence / homogeneity"
output: pdf_document
---

```{r}
belief <- read.table("degreebelief.txt")
belief$Degree <- factor(belief$Degree, levels = c("None", "HighSch", "College"))
(btable <- xtabs(Freq ~ Degree + Belief, belief))
(muhat <- outer(margin.table(btable, 1), margin.table(btable, 2)))/sum(btable) # Takes outer product of vectors giving \hat\pi_{i+}\hat\pi_{+j}
```

# Pearson chi-square test:

```{r}
(X2.test <- chisq.test(btable, correct = F))
```

# LRT

```{r}
(G2 <- 2 * sum(btable * log(btable / muhat)))
1 - pchisq(G2, df = 10)
```

# Residuals

## Pearson

```{r}
X2.test$residuals
```

## Standardized residuals

```{r}
X2.test$stdres
```

# Mosaic plot

```{r}
mosaicplot(btable, shade = T)
```


# Fisher exact test

Test based on hypergeometric distribution

Fixing row totals and column totals (therefore table total is also fixed)

In 2x2 table the degrees of freedom is 1. So we can focus on just 1 element. Say $N_{11}$. Under null hypothesis of independence, $N_{11}$ is conditionally hypergeometric

$$P_{H_0}(N_{11} = t) = \frac{{n_{1+} \choose t} \times {n_{2+} \choose n_{+1}-t}}{{n \choose n_{+1}}}$$

- $H_{a1}: \theta > 1$ for $N_{11}$ larger than expected. $P_{H_0}(N_{11} \ge t_0)$
- $H_{a1}: \theta < 1$ for $N_{11}$ smaller than expected. $P_{H_0}(N_{11} \le t_0)$
- $H_0: \theta = 1$ for $N_{11}$ is as expected

Comments:

- Can use mid p-values
- `fisher.test` in R, can be extended to IxJ tables with some computational difficulties

# Ordinal variables

Departure from independence are monotone: higher levels of X are observed with higher/lower levels of Y

Procedure:

- Assign numerical values (scores) to categories of X and Y (ask an expert)
- Calculate sample correlation
- Reject $H_0$: independence if sample correlation is far away from 0

# Introduction to generalized linear models

Regression model of the form $Y = \beta_0 + \beta_1x_1 + ... + \beta_px_p + \epsilon = E(Y) + \epsilon$ is not appropriate if Y is binary or count.

## Generalized linear models

- Random component: distribution of responses. Assumed to be from natural exponential family of the form $f(y_i;\theta_i) = a(\theta_i)b(y_i)exp(y_iQ(\theta_i))$. $Q(\theta_i))$ is the natural parameter.
- Fixed component: 