---
title: "Example: Residuals of GLM"
output: pdf_document
---

## Example

```{r}
horseshoe <- read.table("horseshoe.txt", header = T)
model <- glm(satell ~ color + spine + width + weight, family = poisson, data = horseshoe)
summary(model)
```

### Leverages

```{r}
plot(hatvalues(model), type = "h")
```

### Standardized residuals

```{r}
plot(rstandard(model, type = "pearson")) # We use standardized residuals of pearson type, not deviance type
```

# Overdispersion

var(Y) larger than expected from fitted values $\hat\mu$. Common in Poisson regression, may occur in binomial regression. Underdispersion may also happen (var(Y) smaller than expected from fitted values $\hat\mu$)

Sources of overdispersion:

- heterogeneity among observations (variations in $\mu$ not captured in the model)
- lurking variables (unused predictors (possibly unknown), higher order terms)
- correlation among observations (eg. clustering)

Under GLM let var(Y) = $\nu^*(\mu)$ when $\mu = E(Y)$

Overdispersion: 

Remedy: quasi-likelihood with additional parameter $\phi$ that models variance separately from mean.

$\nu_\phi(\mu) = \phi \nu^*(\mu)$; For Poisson GLM: $\nu_\phi(\mu) = \phi \mu$; $\phi > 1 \implies$ overdispersion

$\hat\beta$ remains the same under quasi-likelihood. Therefore, $\hat\mu$ remains the same.

$\phi$ is estimated as $\hat\phi = \frac{X^2}{N-p'}$ where $X^2 = \sum_i\frac{(y_i-\hat\mu_i)^2}{\nu^*(\hat\mu_i)}$

Only the standard error of estimates will change: New standard error = $SE(\hat\beta_{j-new}) = \sqrt{\hat\phi}*SE(\hat\beta_{j-old}) \implies$ p-value of 2 sided test will be larger than before.

```{r}
var(rstandard(model, type = "pearson"))
(phihat <- sum(residuals(model, "pearson")^2)/df.residual(model))
summary(model, dispersion = phihat)
```

### Modified LRT

```{r}
drop1(model, test = "F")
```

Shows a warning because model was fit with likelihood, but requested summary uses quasi likelihood. Quasi-Poisson: quasi likelihood with a Poisson model

# Fitting quasipoisson directly

```{r}
quasimodel <- glm(satell ~ width + weight + spine + color, data = horseshoe, family = quasipoisson)
summary(quasimodel)
```

AIC requires a true likelihood, therefore it is NA.

```{r}
drop1(quasimodel, test = "F")
```

No warning!

## Other remedies using true likelihood instead of quasi likelihood:

- Use negative binomial instead of Poisson. It has 2 parameters, therefore it can fit mean and variance separately. It works only for overdispersion.
- Use beta binomial can allow overdispersion in binomial case.

- Another alternative: random effects (mixed models)