---
title: "Test with stratification variable"
output: pdf_document
---

## Example:

```{r}
infectioncure <- read.table("infectioncure.txt", header=TRUE)
head(infectioncure)
# Response coding: 1: success, 2: failure.

# Reshaping for logistic regression: successes and failures for one clinic-treatment pair should be in same row
ic <- reshape(infectioncure, varying=list(c("success","failure")),
              v.names="count", timevar="response",
              idvar=c("center","treat"), direction="wide")

# There are 0 counts => homogeneity must be assumed

# Odds ratio estimates

for(k in unique(ic$center))
  with(ic[ic$center==k,], print(success[treat==1] * failure[treat==2] /
                                  (success[treat==2] * failure[treat==1])))

Treatment <- factor(c("Drug","Control")[ic$treat])
mod <- glm(cbind(success,failure) ~ Treatment + center, family=binomial,data=ic)
summary(mod)

# Wald test suggests significant Treatment (Drug). Center also looks significant
drop1(mod, test="Chisq")
```

**Remarks**

- If we did not account for center, p-value of treatment would have been 0.1 (insignificant).
- Assumes homogeneous association across centers => common conditional odds ratio = exp(0.7769) = 2.2. Transformed Wald interval of conditional odds ratio = $exp(0.7769 \pm 1.96*0.3067) = (1.19, 3.97)$
- Profile likelihood CI of odds ratio: `exp(confint(mod)["TreatmentDrug", ])` = (1.2, 4.02)
- If we didn't account for center, MLE will give odds ratio of 1.5 and confidence interval contains 1.

## Test for homogeneous association

### $G^2$ test

```{r}
1 - pchisq(deviance(mod), df.residual(mod))
```

### Pearson $\chi^2$ test

```{r}
1-pchisq(sum(residuals(mod, type = "pearson")^2), df.residual(mod))
```

$\chi^2$ approximation may not be valid for these tests

## CMH test

```{r}
ic.array <- xtabs(count ~ treat + response + center, data = infectioncure)
ic.array
```

Some of these tables have 0 values

```{r}
mantelhaen.test(ic.array, correct = FALSE)
```

- correct = TRUE does continuity correction
- CI of odds ratio is (1.18, 3.87) - obtained using approximation of log of CMH estimate
- Non-model based method is giving same result as model based method
- CMH is valid even when logistic regression is not valid
- Exact conditional test (generalization of Fisher exact test) if counts are too few that CMH is not valid

# When MLE does not exist

Logistic regression MLE is unique if it exists. Log-likelihood is strictly concave when X has full rank. But it doesn't always exist.

Let $x_i^T$ be ith row of X matrix. Y is binary

MLE exists if there does not exist a non-zero $b \neq 0$ such that $\forall$ i,

- $b^Tx_i > 0 \implies y_i=1$
- $b^Tx_i < 0 \implies y_i=0$

If X has full rank, then existence of b above implies MLE does not exist. There is no maximum.

## Special case: complete separation - MLE doesn't exist

There exists b such that:

- $y_i = 1 \implies b^Tx_i > 0$
- $y_i = 0 \implies b^Tx_i < 0$

Comments:

- If there is no intercept, there exists a subspace in column space of X that perfectly separates the classes
- If there is intercept, there exists a hyperplane that perfectly separates classes
- Likelihood converges to 1