---
title: "Session 3"
author: "Naveen Mathew Nathan S."
output:
  pdf_document: default
  word_document: default
  html_document: default
---
# Session 3

`optim` function can be used to optimize a known funcitonal form of parameters.

Parameters should be the first parameter of the function that is passed to `optim`

## Linear Regression

```{r, eval = FALSE, echo = TRUE}
# eval = FALSE only generates the code without running it
# echo = FALSE removes the code chunk from the report itself!
loss <- function(b, X, Y) {
  return(((Y - X %*% b)^2)/length(Y))
}
```

Scope of a variable in a function is only within the funciton. Global values will not be changed unless we use <<-

```{r}
a = 1:10

changea <- function(a) {
  a <- a + 1
  return (a)
}

changea(a)
a

changea1 <- function(a) {
  a <<- a + 1
  return (a)
}

changea1(a)
a

changea2 <- function(env) {
  env$a <- env$a + 1
  return (env$a)
}

changea2(env = environment())
a

changea3 <- function(b) {
  a <- a+1
  return(a)
}

changea3(b = 1:5) # Picks up 'a' from the previous level environment and computes the function
```

- To list all objects in current environment: `ls()`
- To remove an object from current environment: `rm('variablename')`

- `foreach` package creates a copy of the environment in each worker node