---
title: "STAT429 - Session 1"
author: "Naveen Mathew Nathan S."
date: "8/26/2019"
output: pdf_document
---

## Examples

### Johnson & Johnson stock price

```{r}
library(astsa)
plot(jj, type = 'o')
```

General pattern:

- increasing trend
- sharper increase/decrease in later years

### Global temperature deviations

```{r}
plot(globtemp)
abline(h = 0)
```

Shows ups and downs but increasing in general.

```{r}
plot(nyse)
```

Mostly stable, high volatility in a short period.

```{r}
plot(speech)
```

Patterns repeat.

```{r}
library(xts)
djiar <- diff(djia$Close)/djia$Close[-nrow(djia)]
plot(djia, col = "black")
```

Stochastic volatility models may be required to model the variation here.

## Time Series

- Data collected over time.
- Time oriented or chronological.
- Sequence of observations on a variable of interest.
- Data cannot be assumed independent.
- Time series models incorporate time dependence.

### Approaches

#### Time domain approach

Modeling future value of a time series based on a parametric function of current and past values.

#### Frequency domain approach

Assume the primary feature of a time series is periodical or systematic sinusoidal variations. These variations are naturally found in most data sets.

Both approaches produce similar results on long time series. Ideally - use both in complementary way.

## Time domain approach:

### Box and Jenkins approach: ARIMA, transfer function

- Additive model: observed data arise from the sum of structured components. Eg: $trend + seasonal + effect error$
- Multiplicative model: Observerd data arise from multiplicative model involving product of factors.

## Three steps for model building

- Model specification
- Model fitting
- Model diagnostics

# Problem set:

1.2 a) and b)